---
title: "Soil drivers"
author: "Romy Zeiss"
date: "`r format(Sys.time(), '%B %Y')`"
header-includes:
- \usepackage{pdflscape} 
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../doc", output_format = "all") })
fontsize: 8pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(formattable)
```

## Literature review on drivers of soil biota

First, we made a literature review on the drivers of different soil organisms. The results of this review can be found in the following table.

```{r}
setwd("I:/eie/==PERSONAL/RZ SoilBON/SoilBiodiversity")
Env_Predictors_background <- read_csv("doc/Env_Predictors_background.csv")
head(Env_Predictors_background)
```

Now, we will make this table a bit more nice, and combine similar relationships.

```{r}

Tab1 <- Env_Predictors_background %>% group_by(Category, Response) %>% 
  mutate_all(~str_replace_na(., "")) %>%
  mutate(values=paste(Reference, ": ", Predictor, " ", Effect_size, " ", Variance)) %>%
  summarise_all(paste, collapse="/ ", na.action(na.omit))  %>%
  select(Category, values, Response) %>%
  pivot_wider(names_from=Response, values_from=values)
```

\newpage
\blandscape

```{r}
options(knitr.kable.NA = "")
knitr::kable(Tab1, caption = "Table 1. Drivers of soil biodiversity.")
```
\elandscape

### Müller et al. 2016

Here now separately the results of one extensive study.

```{r}
Tab2 <- Env_Predictors_background %>% 
  filter(., Reference=="Mueller et al. 2016") %>%
  group_by(Category, Response) %>% 
  mutate_all(~str_replace_na(., "")) %>%
  select(Category, Variance, Response, Long_name, Predictor) %>%
  pivot_wider(names_from=Response, values_from=Variance)
```

\newpage
\blandscape

```{r echo=T} 
# unfortunately, this doesn't work in PDF...

formattable::formattable(Tab2, align =c("l","l","l","c","c", "c", "c", "c", "c", "c"), list(
  `Category` = formatter("span", style = ~ style(color = "grey",font.weight = "bold")),
  `Nematods`= color_bar("#DeF7E9"),
  `Earthworms`= color_bar("#DeF7E9"),
  `Mites`= color_bar("#DeF7E9"),
  `Collembola`= color_bar("#DeF7E9"),
  `Ants`=color_bar("#DeF7E9"),
  `Invertebrates`= color_bar("#DeF7E9")))
```

```{r}
Tab2 <- Env_Predictors_background %>% 
  filter(., Reference=="Mueller et al. 2016") %>%
  group_by(Predictor, Response, Category) %>% 
  summarize(mean_Var = mean(Variance)) %>%
  ungroup() %>% 
  mutate_all(~str_replace_na(., "")) %>%
  select(Category, mean_Var, Predictor, Response) %>%
  pivot_wider(names_from=Response, values_from=mean_Var) %>%
  arrange(Category)

# create numeric matrix
Mat2 <- matrix(
    as.numeric(
      as.matrix(Tab2[,3:8], ), 
      ncol = 6), 
    ncol=6, dimnames=list(Tab2$Predictor, colnames(Tab2[,3:8])))

# define colors
my_group <- as.numeric(as.factor(substr(Tab2$Category, 1 , 1)))
colSide <- RColorBrewer::brewer.pal(4, "Set1")[my_group]

mydf <- data.frame(row.names = rownames(Mat2), category = Tab2$Category)

heatmap(Mat2, na.rm=T, Colv = NA, Rowv = NA, RowSideColors=colSide)

pheatmap::pheatmap(Mat2 ,  na.rm=T, cluster_cols = F, cluster_rows = F, 
                   annotation_row = mydf, gaps_row = c(5, 6, 15), cellheight = 10, cellwidth = 20)
```

\elandscape


